Changes in Patterns of Geometric Sign Groupings as a Model of Social Dynamics Across Aurignacian Phases

### 2025-05-06 Extra/undergraduate-research-symposium-images.qmd
edit

>>>>

```{r}
sign_type_df <- signbase_full_clean %>% 
  pivot_longer(cols = line:star) %>% 
  filter(value != 0) %>% 
  mutate(sign_type = name) %>% 
  group_by(sign_type) %>% 
  summarise(value = n()) %>% 
  arrange(desc(value))

sign_type_plot <- ggplot(sign_type_df) +
  aes(reorder(sign_type, value), x = sign_type, y = value) +
  geom_col() +
  xlab("Sign type") +
  ylab("Count") +
  theme(axis.title = element_text(size = 10)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, ))

ggsave(sign_type_plot, 
      filename =  "signtype.png",
       bg = "white", w = 5, h = 3)

```{r}
library(rnaturalearth)

as_factor <- function(x, reverse = FALSE) {
  lvl <- unique(x)
  if (reverse) {
    lvl <- rev(lvl)
  }
  factor(x, levels = lvl)
}

to_long <- function(from, factor = FALSE, reverse = FALSE) {
  x <- data.frame(
    row = as.vector(row(from, as.factor = factor)),
    column = as.vector(col(from, as.factor = factor)),
    value = as.vector(from),
    stringsAsFactors = FALSE
  )
  if (factor) {
    x$row <- as_factor(x$row, reverse = reverse)
    x$column <- as_factor(x$column, reverse = reverse)
  }
  return(x)
}

autoplot.DiversityIndex <- function(object, ...) {
  ## Prepare data
  count <- cbind.data.frame(
    label = object@labels,
    x = object@size,
    y = object@.Data
  )

  ## Simulated assemblages
  gg_sim <- NULL
  if (length(object@simulation) != 0) {
    # Build a long table for ggplot2
    refined <- object@simulation

    sim_stacked <- to_long(refined[, -c(1)], factor = TRUE)
    sim <- cbind.data.frame(
      size = refined[, 1],
      sim_stacked,
      Estimate = ifelse(sim_stacked[["column"]] == "mean", "mean", "conf. int.")
    )
    gg_sim <- ggplot2::geom_path(
      mapping = ggplot2::aes(
        x = .data$size,
        y = .data$value,
        colour = .data$Estimate,
        group = .data$column
      ),
      data = sim,
      na.rm = TRUE,
      inherit.aes = FALSE
    )
  }

  y_lab <- switch (
    class(object),
    HeterogeneityIndex = "Heterogeneity",
    EvennessIndex = "Evenness",
    RichnessIndex = "Richness",
    "Diversity"
  )

  ## ggplot
  ggplot2::ggplot(data = count) +
    ggplot2::aes(x = .data$x, y = .data$y, label = .data$label) +
    ggplot2::geom_point() +
    gg_sim +
    ggplot2::scale_x_log10(name = "Sample size") +
    ggplot2::scale_y_continuous(
      name = "Heterogeneity"
    )
}

plot_diversity_fn <- function(input_df){

signs <-
  input_df %>% 
  select(
         site_name,
         c(line:star)) %>% 
  group_by(site_name) %>% 
  summarise(across(where(is.numeric), ~ sum(.))) %>% 
  column_to_rownames("site_name")

diversity_index <- 
  heterogeneity(signs, method = "shannon")

diversity_sim <- simulate(diversity_index)

autoplot.DiversityIndex(diversity_sim) +
  geom_text_repel(aes(label = diversity_sim@labels),
                  size = 3, max.overlaps = 15) +
  guides(colour = "none") +
  theme_minimal(base_size = 8) +
  border(color = "black", size = 0.5)
}

signbase_full_clean_diversity_plot_by_period <- 
time_unique_data_full %>% 
  nest_by(time_period) %>% 
  mutate(plot = list(plot_diversity_fn(data)))

library(cowplot)
plot_grid(plotlist =  signbase_full_clean_diversity_plot_by_period$plot,
          labels = signbase_full_clean_diversity_plot_by_period$time_period,
          label_size = 8, vjust = 2, hjust = -0.6)
```
